---
title: "TidyModels Playaround"
author: "Leonard Seok"
date: "24 May 2020"
output: 
  html_document:
    theme: journal
    hightlight: tango
---

Lets load the PimaIndiansDiabetes dataset for this play-around

```{r setup, results = 'hide'}
# A custom function to install and load packages
install.custom <- function(package.name) {
  if (!(package.name %in% rownames(installed.packages()))) {
    install.packages(package.name)
  } else {
    print(paste0(package.name," ","is already installed"))
  }
  library(package.name, character.only = T)
}

install.custom("purrr")
list_package <- list("mlbench", "dplyr", "tidyverse", "ggplot2")
map(list_package, ~install.custom(.x))

data("PimaIndiansDiabetes") # Load the data 

data <- PimaIndiansDiabetes
```

Investigate the Numeric data with some histograms

```{r}
data %>% dplyr::select_if(is.numeric) %>% 
  gather(key='field', value='value') %>%
  ggplot(aes(x=value)) + geom_histogram(color='black') +
  facet_wrap(~field, scales = 'free')
```

We can see that for some fields such as insulin, pregnant, and pressure has a large concentration of 0. 
It doesn't quite makes sense that mass and tricep skin fold are recorded as 0, so they are most probably missing values.

```{r}
data %>% ggplot(aes(x=diabetes)) + geom_bar()
```

For all fields other than pregnant and diabetes, lets set 0 to be NA

```{r}
d1 <- data %>% select(-pregnant,-diabetes) %>% 
  map_dfr(~ifelse(.x==0,NA,.x))

data_clean <- cbind(d1, data %>% select(pregnant, diabetes))

data_clean %>% select_if(is.numeric) %>% 
  gather(key='key', value = 'value') %>% 
  ggplot(aes(x=value)) + geom_histogram(color='black', fill='cyan') +
  facet_wrap(~key, scales = 'free')
```

## Modelling


